---
title: "Installation Guide for rSCHNEL"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installation Guide for rSCHNEL}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this tutorial tutorial we will be showing how to prepare the rSCHNEL package in order to support the full functionality of the package for many different input types. 

In order to ensure the successful installation of the package, download two libraries that the wrapped C++ HSNE dimensionality reduction implementation requires. 
In the terminal, run:
```
$ sudo apt-get update
$ sudo apt-get install liblz4-dev
$ sudo apt-get install libflann-dev
```

To support the use of the clustering algorithm, please install python modules for the leiden algorithm.
In case of any problems refer to the official website [here](https://cran.r-project.org/web/packages/leiden/vignettes/run_leiden.html)
Install the required packages with `pip install leidenalg` / `pip 3 install leidenalg`

To download the rSCHNEL package, first install `devtools` utilities package
```{r, eval = F}
install.packages("devtools")
```

Then, download the package from GitHub.
```{r, eval = F}
devtools::install_github(biovault/rschnel)
```

To support different input types the rSCHNEL package makes use of open-source tools available from different sources.
Here we present all the required packages for each file/object type.

## `.fcs` files
In order to use the `.fcs` files install `flowCore` package from bioconductor 
```{r, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("flowCore")
```
To support parsing the flowFrame objects install the ParkerICI/grappolo project
```{r, eval=FALSE}
devtools::install_github("ParkerICI/grappolo")
```

That allows you to read in and cluster data from `.fcs` files
```{r, eval=FALSE}
clusterSingleFile <- cluster(“../../data/fcs_data.fcs”)
clusterDirectory <- cluster(“../../data/fileFCS/”, dir = T)
```

## `Seurat objects`

The `Seurat` package should be automatically installed when downloading the rSCHNEL package. 
In case of unsuccesful installation, install the package yourself by running:
```{r, eval = FALSE}
install.packages("Seurat")
```

With help of this library, the rSCHNEL package supports operations on Seurat objects.

```{r, eval=FALSE} 
# Extract the data from the Seurat object into a matrix
seurat_to_matrix <- parse_to_matrix("../../data/seuratObj.rds")

# Perform clustering on a seurat object
cluster_seurat <- cluster("../../data/seuratObj.rds")
```

## `.h5ad files`
Parsing `.h5ad` files in the rSCHNEL package makes use of two packages (currently under development) available on gitHub. 
These packages are SeuratDisk and SeuratData. They allow for parsing `.h5ad` files into a Seurat object which is parsed within the package and fed into the clustering algorithm. 

To install the required packages run the following commands:
```{r, eval=FALSE}
devtools::install_github("mojaveazure/seurat-disk")
devtools::install_github('satijalab/seurat-data')
```


## SingleCellExperiment objects
The rSCHNEL package can take as input SingleCellExperiment objects
 Parsing of the SingleCellExperiment first transforms the SCE object into a Seurat object. To support this feature install a package from bioconductor.
```{r, eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("SingleCellExperiment")
```

